Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-09-20 paper.tex
Added bibliography tex file, minor changes in Related works

\begin{abstract}
 Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.  Heterogeneous  re- sources are distributed worldwide, thousands of physicists analyz- ing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year. The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment. In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model. The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.  PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
2013. We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades. Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers. Thousands of physicists analyze peta-scale  data volumes daily. One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis. The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources. Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
PanDA has a highly scalable architecture. Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade. PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware). This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.  Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
\end{center}
\end{figure}

@@

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
\input{WMSLiterature}
--------------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

@@

\subsection{PanDA Pilot}
 PanDA is a pilot based workload  management system. In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites. When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it. Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server. First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.  The repository works with Apache web server configured  with a  memory-based   web proxy (Squid). The purpose  of the cache  is to reduce  the

request load on the back-end Subversion  server. That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server. The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.

\subsection{PanDA Factory}
As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites. This is done with the help of a component  called AutoPyFactory (APF). APF runs in a single daemonized process, launching  a separate thread for each internal  workflow. Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote. The behavior of these APF workflows is determined by the combination of a set  of plug-ins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.

\subsection{PanDA Monitoring}
The PanDA Monitor is a web-based dashboard-style  graphi- cal application. It runs on Apache and interacts with the back- end database for persistence. The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory. The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA develop- ers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6]. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}

@@

 \section{Conclusion}\label{sec:conclusion}
The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of
computing in HEP was  discarded  in favor  of a  far more flexible and scalable  model. The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments. PanDA  is the first exascale  workload manage- ment system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013. Next LHC run will pose massive computing  challenges. With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more. Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS. To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system. Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry. The technology base provided by the PanDA system will enhance the usage of a variety  of high- performance computing resources available to basic research.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

\begin{abstract}
 Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.  Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year. The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment. In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model. The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.  PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
2013. We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades. Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers. Thousands of physicists analyze petascale  data volumes daily. One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis. The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources. Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
PanDA has a highly scalable architecture. Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade. PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware). This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.  Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
\end{center}
\end{figure}

@@

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
\input{WMSLiterature}
--------------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

@@

\subsection{PanDA Pilot}
 PanDA is a pilot based workload  management system. In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites. When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it. Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server. First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.  The repository works with Apache web server configured  with a  memory-based   web proxy (Squid). The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server. That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server. The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.

\subsection{PanDA Factory}
As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites. This is done with the help of a component  called AutoPyFactory (APF). APF runs in a single daemonized process, launching  a separate thread for each internal  workflow. Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote. The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.

\subsection{PanDA Monitoring}
The PanDA Monitor is a web-based dashboard-style  graphical application. It runs on Apache and interacts with the back-end database for persistence. The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory. The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA developers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}

@@

 \section{Conclusion}\label{sec:conclusion}
The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model. The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments. PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013. Next LHC run will pose massive computing  challenges. With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more. Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS. To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system. Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry. The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

\bibliographystyle{plain}
\bibliography{bibliography}
